---
title: "Group 4 — Youth risk"
subtitle: "Studio wall"
author: "Shivgun, Shreeya and Meghana"
format:
  html:
    toc: true
    toc-depth: 2
---

# Spine

> What do teenagers’ bodies and habits look like, which habits travel together, and where is life long?

This sentence stays at the top. Every panel must attach to it. If a plot cannot be introduced in a clause that hangs off this line, it is off the wall.

# Introduction — the tables we own

This folder is ours for the ten days. Three tables, three shapes.

- **Panel A** (`panel-A.csv`) — YRBSS Quants: height, weight, days active, strength training, hours of sleep.
- **Panel B** (`panel-B.csv`) — YRBSS Quals: helmet, texting while driving, TV, race, gender, grade. Same survey, different columns.
- **Panel C** (`panel-C.csv`) — gapminder 2007 world. Ask a *different* map question than Group 3 (wealth, population symbols, continent).

A and B are the same teenagers cut two ways. Do not mix them into one dishonest figure.

We load from this folder:

```{r}
#| label: setup
library(tidyverse)
library(skimr)
library(mosaic)
library(readr)

PanelA <- read.csv(file = "data/panel-A.csv")
PanelB <- read.csv(file = "data/panel-B.csv")
PanelC <- read.csv(file = "data/panel-C.csv")

```

Found tables train the mark. A campus Free Hunch, if we collect one, is a **made** panel of the same pair type — new people, old grammar. It does not replace A, B, or C.

# Panel A — body / measure

*height, weight, physically_active_7d, school_night_hours_sleep are Quant. grade is Qual.*




### 1. See the table

What is a case? Which columns are Quant / Qual? What is missing?

```r
# glimpse() / inspect() / skim()
```

```{r}
#| label: inspect-panelA-data-1
dplyr::glimpse(PanelA)

```
My PanelA data has:

  - 13,583 observations and 08 variables.
  - `age`: Quantitative, numeric
  - `height`: Quantitative, numeric
  - `weight`: Quantitative, numeric
  - `physically_active_7d`: Quantitative, numeric
  - `strength_training_7d`: Quantitative, numeric
  - `gender`: Qualitative, categorical. Encoded <chr>, to be mutated
  - `grade`: Qualitative, categorical. Encoded <chr>, to be mutated
  - `school_night_hours_sleep`: Qualitative, categorical. Encoded <chr>, to be mutated

```{r}
#| label: inspect-panelA-data-2
skimr::skim(PanelA)
```

Furthermore, on the Qual variables panel-A data has : 

- `gender` has 3 levels: 
- `grade` has 6 levels: 
- `school_night_hours_sleep` has 8 levels: 

And on the Quant variables panel-A data has :

- `age` has a mean of 16.15, sd of 1.26, median of 16.00, min of 12.00, and max of 18.00
- `height` has a mean of 1.69, sd of 0.10, median of 1.68, min of 1.27, and max of 2.11
- `weight` has a mean of 67.90, sd of 16.89, median of 64.41, min of 29.94, and max of 180.99
- `physically_active_7d` has a mean of 3.90, sd of 2.56, median of 4.00, min of 0.00, and max of 7.00
- `strength_training_7d` has a mean of 2.94, sd of 2.57, median of 3.00, min of 0.00, and max of 7.00

There is missing data.

```{r}
#| label: mutate-panelA-factors

A_mod <- PanelA |>  mutate( gender = as.factor(gender),
    grade = as.factor(grade),
    school_night_hours_sleep = as.factor(school_night_hours_sleep))

glimpse(A_mod)
    
```
All is well now.


### 2. Ask

One question *this* table can answer. One it cannot. Write both.

Looking at the overall Spine statement, I can pose the following Questions:

1. Which grade has the most students?
2. What is the sex distribution among the grades?
2. Do boys strength train more than girls?



### 3. Choose a mark

Geom, map mark, or (later) test that matches the pair type. Name it before you code.

histogram coloured by gender 


### 4. Predict

What do we think we will see? One sentence, before Run.
the counts of boys who strength train might be higher.

### 5. Make

Human brick first. Optional: one AI prompt, pasted below, only for what section 3 named.

1. Which grade has the most students??

```{r}
A_mod %>% gf_bar (~ grade)
```
2. What is the sex distribution among the grades?

```{r}
A_mod %>% gf_bar(~grade, fill = ~gender)
```



3. Do boys strength train more than girls?

```{r}
A_mod %>% gf_histogram(~ strength_training_7d, fill = ~gender, bins = 1, position = "dodge")
```
4. Do girls have better sleep than boys?

```{r}
A_mod %>% gf_bar(~ gender, fill = ~school_night_hours_sleep)
```

5. 

```r
# human chunk

# ## Prompt we actually used
# ai_draft
# ours
```


### 6. Say

Caption + one sentence we will stand behind. Title is a sentence, not a column name.

### 7. Limit

What this panel cannot do. Which kit table we refused to abuse.


# Panel B — category / talk

*helmet_12m × gender, text_while_driving_30d × grade. Hours of TV may be ordinal — argue before you average it.*


### 1. See the table

What is a case? Which columns are Quant / Qual? What is missing?

```r
# glimpse() / inspect() / skim()
```
```{r}
#| label: inspect-panelB-data-1
dplyr::glimpse(PanelB)

```
My PanelB data has:

  - 13,583 observations and 08 variables.
  - `age`: Quantitative, numeric
  - `gender`: Qualitative, categorical. Encoded <chr>, to be mutated
  - `grade`: Qualitative, categorical. Encoded <chr>, to be mutated
  - `hispanic`: Qualitative, categorical. Encoded <chr>, to be mutated
  - `race``: Qualitative, categorical. Encoded <chr>, to be mutated
  - `helmet_12m`: Qualitative, categorical. Encoded <chr>, to be mutated
  - `text_while_driving_30d`: Qualitative, categorical. Encoded <chr>, to be mutated'
  - `hours_tv_per_school_day`: Qualitative, categorical. Encoded <chr>, to be mutated

```{r}
#| label: inspect-panelB-data-2
skimr::skim(PanelB)
```
Furthermore, on the Qual variables panel-A data has : 

- `gender` has 3 levels: 
- `grade` has 6 levels: 
- `hispanic` has 3 levels: 
- `race`has 6 levels:
- `helment_12m` has 7 levels:
- `text_while_driving_30d` has 9 levels:
- `hours_tv_per_school_day` has 8 levels:

And on the Quant variables panel-A data has :

- `age` has a mean of 16.15, sd of 1.26, median of 16.00, min of 12.00, and max of 18.00

```{r}
#| label: mutate-panelA-factors

B_mod <- PanelB |>  mutate( gender = as.factor(gender),
    grade = as.factor(grade), hispanic = as.factor(hispanic), race = as.factor(race), helmet_12m = as.factor(helmet_12m), text_while_driving_30d = as.factor(text_while_driving_30d), hours_tv_per_school_day = as.factor(hours_tv_per_school_day)) 

glimpse(B_mod)
```


### 2. Ask

One question *this* table can answer. One it cannot. Write both.

### 3. Choose a mark

Geom, map mark, or (later) test that matches the pair type. Name it before you code.

```{r}
vcd::structable(hours_tv_per_school_day ~ gender, data = B_mod) %>% vcd::mosaic()
```


### 4. Predict

What do we think we will see? One sentence, before Run.

### 5. Make

Human brick first. Optional: one AI prompt, pasted below, only for what section 3 named.

```r
# human chunk

# ## Prompt we actually used
# ai_draft
# ours
```

### 6. Say

Caption + one sentence we will stand behind. Title is a sentence, not a column name.

### 7. Limit

What this panel cannot do. Which kit table we refused to abuse.


# Panel C — place

*Same file as Group 3 C. Different question. lifeExp vs gdpPercap, or pop as symbol size.*

Geometry is not in the CSV. Join in R, then use the same seven headings.

```r
# library(sf)
# library(rnaturalearth)
# world <- ne_countries(scale = "medium", returnclass = "sf")
# # join on country name; list the unmatched labels
```


### 1. See the table

What is a case? Which columns are Quant / Qual? What is missing?

```r
# glimpse() / inspect() / skim()
```
```{r}
#| label: inspect-panelC-data
dplyr::glimpse(PanelC)

```


### 2. Ask

One question *this* table can answer. One it cannot. Write both.

### 3. Choose a mark

Geom, map mark, or (later) test that matches the pair type. Name it before you code.

### 4. Predict

What do we think we will see? One sentence, before Run.

### 5. Make

Human brick first. Optional: one AI prompt, pasted below, only for what section 3 named.

```r
# human chunk

# ## Prompt we actually used
# ai_draft
# ours
```

### 6. Say

Caption + one sentence we will stand behind. Title is a sentence, not a column name.

### 7. Limit

What this panel cannot do. Which kit table we refused to abuse.


# Conclusion

What the spine looks like after three panels.

- The question we will defend in ninety seconds
- What we will **not** claim
- Which lines were human, which were machine
- Limits: sample, missingness, the test we did not run
- Gift for the steal-this strip (palette, annotation, join trick)

```r
# sessionInfo()
```
